The Calculus of Linear Constructions

### 2021-10-12 writing/draft1.tex
draft

      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{7.3em} expressions
      \indalt $(x : A) \arw{s}{t} B$
      \indalt $\lambda x. n$ | $m\ n$
    \end{grammar}
    \end{minipage}
    \label{syntax}
  \end{figure}
  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory is the indexed function type: $(x : A) \arw{s}{t} B$. The reason for these indices is that we have built the ! exponential of linear logic directly into universe sorts. The indices $s$, and $t$ annotate the domain, and co-domain's respective universe sort, hence annotating linearity. The behavior of ! is difficult to account for even in simple linear type theory. Subtle issues arise if !! is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}. By integrating the exponential directly into universe sorts, we implicitly limit ! to only be used canonically. This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without needing any machinery for manipulating exponential.
  \begin{figure}[H]
    \caption{Correspondence of CLC types and MELL implications}
    \begin{align}
      (\_ : A) \arw{U}{U} B \quad &\equiv \quad !A \multimap !B \\
      (\_ : A) \arw{U}{L} B \quad &\equiv \quad !A \multimap B \\
      (\_ : A) \arw{L}{U} B \quad &\equiv \quad A \multimap !B \\
      (\_ : A) \arw{L}{L} B \quad &\equiv \quad A \multimap B
    \end{align}
    \label{correspondence}
  \end{figure}
  Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications. MELL lacks counterparts for the cases (1), (2) if the co-domain $B$ is dependent on arguments of domain $A$. Though it may seem as if implications of the $!(A \multimap B)$ form are left out, CLC encodes these by endowing each function type with a universe sort through their formation judgment. The encoding for this example would be $\Gamma \vdash A \arw{L}{L} B \utype U_i $. We will discuss function type formation in Section \ref{tyformation} in greater detail.

@@

  \subsection{Equality and Parallel Reduction} \label{reduction}
  The operational semantics, and definitional equality of CLC is defined by parallel reductions \cite{takahashi}. The reduction, and equality rules for CLC are presented in Figure \ref{parallel}, and are entirely standard.
  \begin{figure}[H]
    \caption{Equality and Parallel Reduction}
    \begin{mathpar}
      \inferrule
      { m_1 \pstep^* n \\ m_2 \pstep^* n }
      { m_1 \equiv m_2 : A }
      \rname{Join}

@@

  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions. This is contradictory to the strong normalization property of CC$\omega$, thus this hypothetical term does not exist in CLC.

>>>>

      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{7.3em} expressions
      \indalt $(x : A) \arw{s}{t} B$
      \indalt $\lambda x. n$ | $m\ n$
    \end{grammar}
    \end{minipage}
    \label{syntax}
  \end{figure}
  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory is the indexed function type: $(x : A) \arw{s}{t} B$. The reason for these indices is that we have built the ! exponential of linear logic directly into universe sorts. The indices $s$, and $t$ annotate the domain, and co-domain's respective universe sort, hence annotating linearity. The behavior of ! is difficult to account for even in simple linear type theory. Subtle issues arise if !! is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}. By integrating the exponential directly into universe sorts, we implicitly limit ! to only be used canonically. This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without needing any machinery for manipulating exponential.
  \begin{figure}[H]
    \caption{Correspondence of CLC types and MELL implications}
    \begin{align}
      (\_ : A) \arw{U}{U} B \quad &\Leftrightarrow \quad !A \multimap !B \\
      (\_ : A) \arw{U}{L} B \quad &\Leftrightarrow \quad !A \multimap B \\
      (\_ : A) \arw{L}{U} B \quad &\Leftrightarrow \quad A \multimap !B \\
      (\_ : A) \arw{L}{L} B \quad &\Leftrightarrow \quad A \multimap B
    \end{align}
    \label{correspondence}
  \end{figure}
  Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications. MELL lacks counterparts for the cases (1), (2) if the co-domain $B$ is dependent on arguments of domain $A$. Though it may seem as if implications of the $!(A \multimap B)$ form are left out, CLC encodes these by endowing each function type with a universe sort through their formation judgment. The encoding for this example would be $\Gamma \vdash A \arw{L}{L} B \utype U_i $. We will discuss function type formation in Section \ref{tyformation} in greater detail.

@@

  \subsection{Equality and Parallel Reduction} \label{reduction}
  The operational semantics, and definitional equality of CLC is defined by parallel reductions \cite{takahashi} as presented in Figure \ref{parallel}, all of which are entirely standard.
  \begin{figure}[H]
    \caption{Equality and Parallel Reduction}
    \begin{mathpar}
      \inferrule
      { m_1 \pstep^* n \\ m_2 \pstep^* n }
      { m_1 \equiv m_2 : A }
      \rname{Join}

@@

  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions. This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
